Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-11 manuscript/manuscript.tex
Update grid table
Small improvements to introduction

Two recent papers present other approaches for homogeneous tesseroids gravitational
fields computation and incorporate methodologies for variable density tesseroid with
specific density profiles.
\citet{Fukushima2018} analytically integrates the volumetric integral for the
gravitational potential in the radial direction, obtaining a surface integral, which is
then numerically solved by applying a conditionally splitting and the double exponential
quadrature rule. The gradient of the potential and the Marussi tensor components are
then computed by finite differences. \citet{Fukushima2018} also generalizes his method
to an arbitrary polynomial density profile on the radial coordinate.
\citet{Lin2018} made a comparison of the different methodologies for homogeneous
tesseroid gravitational fields computation by applying different discretizations
algorithms to each one of them and developed a combined method.
For computation points near the tesseroid they apply a GLQ numerical integration with
a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
dimension. But, if the computation point is farther than a certain truncation distance,
then a second order Taylor series approximation is applied along with the regular
subdivision developed by \citet{Grombein2013}.
They also introduce a variation of their combined method to compute the gravitational
fields generated by tesseroids with linear density on the radial coordinate.

@@

We present a new algorithm for computing the gravitational fields generated by any
tesseroid with any arbitrary continuous density function on any external point.
It is based on the GLQ approximation and a two dimensional version of the adaptive
discretization of \citet{Uieda2016}, which we extend to include a new density-based
discretization algorithm.
In order to ensure the accuracy of the numerical approximation we have
determined optimal values for the controlling parameters by
comparing the numerical approximation with analytical solutions for
spherical shells.
Finally, we applied the methodology to model the Neuqu\'en basin, Argentina, using
tesseroids with an exponentially increasing density with depth.

@@

\begin{table}
\caption{
    Description of the computation grids used to characterize the accuracy of the
    numerical integration.
    All grids consist of a set of $10 \times 10$ points.
}
\label{tab:grids}
\begin{tabular}{lccc}
    Name & Grid size & Grid region (degrees) & Grid height (km)
    \\ \hline
    Pole & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
    Equator & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
    Satellite & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
    Big Grid & $30^\circ \times 30^\circ$ & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
\end{tabular}
\end{table}

>>>>

Two recent papers present other approaches for homogeneous tesseroids gravitational
fields computation and incorporate methodologies for variable density tesseroid with
specific density profiles.
\citet{Fukushima2018} analytically integrates the volumetric integral for the
gravitational potential in the radial direction, obtaining a surface integral, which is
then numerically solved by applying a conditionally splitting and the double exponential
quadrature rule. The gradient of the potential and the Marussi tensor components are
then computed by finite differences. \citet{Fukushima2018} also generalizes his method
to an arbitrary polynomial density profile on the radial coordinate.
\citet{Lin2018} made a comparison of the different methodologies for homogeneous
tesseroid gravitational fields computation by applying different discretizations
algorithms to each one of them.
From this analysis they develop a combined method:
for computation points near the tesseroid they apply a GLQ numerical integration with
a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
dimension. But, if the computation point is farther than a certain truncation distance,
then a second order Taylor series approximation is applied along with the regular
subdivision developed by \citet{Grombein2013}.
They also introduce a variation of their combined method to compute the gravitational
fields generated by tesseroids with linear density on the radial coordinate.

@@

We present a new algorithm for computing the gravitational fields generated by any
tesseroid with any arbitrary continuous density function on any external point.
It is based on the three dimensional GLQ approximation, a two dimensional version of the
adaptive discretization of \citet{Uieda2016} and a new density-based discretization
algorithm that we introduce in this work.
In order to ensure the accuracy of the numerical approximation we have
determined optimal values for the controlling parameters by
comparing the numerical approximation with analytical solutions for
spherical shells.
Finally, we applied the methodology to model the Neuqu\'en basin, Argentina, using
tesseroids with linear and exponentially increasing density with depth.

@@

\begin{table}
\caption{
    Description of the computation grids used to characterize the accuracy of the
    numerical integration.
    Grid height is defined above the mean Earth radius.
}
\label{tab:grids}
\begin{tabular}{lccc}
    Name & Grid size & Grid region (degrees) & Grid height (km)
    \\ \hline
    Pole      & $0.1^\circ \times 0.1^\circ$ & 0E / 1E / 89N / 90N  & 0   \\
    Equator   & $0.1^\circ \times 0.1^\circ$ & 0E / 1E / 0N  / 1N   & 0   \\
    Global    & $ 10^\circ \times  10^\circ$ & 0E / 1E / 0N  / 1N   & 0   \\
    Satellite & $ 10^\circ \times  10^\circ$ & 0E / 1E / 89N / 90N  & 260 \\
\end{tabular}
\end{table}
